\chapter{Methodology}
\label{chap:methodology}

\todo{Write full methodology chapter.}

\section{Problem formulation}

Let $\mathcal{U} = \{u_1, \dots, u_M\}$ and $\mathcal{I} = \{i_1, \dots, i_N\}$
denote the sets of users and POIs respectively.
We observe implicit feedback $\mat{R} \in \{0,1\}^{M \times N}$, where
$r_{ui} = 1$ if user $u$ has interacted with POI $i$.
The goal is to learn a ranking function $\hat{r}_{ui} : \mathcal{U} \times
\mathcal{I} \to \mathbb{R}$ such that unobserved but relevant POIs receive
higher scores than irrelevant ones.

\section{Baseline: matrix factorisation}

We factorise $\mat{R} \approx \mat{P} \mat{Q}^\top$ where
$\mat{P} \in \mathbb{R}^{M \times d}$ and $\mat{Q} \in \mathbb{R}^{N \times d}$
are user and item embedding matrices, trained with Bayesian Personalised
Ranking (BPR) loss.

\section{Sparse Autoencoder on latent factors}

Given a trained user embedding $\vect{p}_u \in \mathbb{R}^d$, a SAE maps it to
a sparse code $\vect{z}_u \in \mathbb{R}^k$ ($k \gg d$):

\begin{align}
  \vect{z}_u &= \sigma\!\left(\mat{W}_{\text{enc}}\,\vect{p}_u + \vect{b}_{\text{enc}}\right), \\
  \hat{\vect{p}}_u &= \mat{W}_{\text{dec}}\,\vect{z}_u + \vect{b}_{\text{dec}},
\end{align}

where $\sigma$ is a ReLU activation that promotes non-negativity and sparsity.
The reconstruction loss is:

\begin{equation}
  \mathcal{L}_{\text{SAE}} = \|\vect{p}_u - \hat{\vect{p}}_u\|_2^2
                            + \lambda \|\vect{z}_u\|_1.
\end{equation}

\section{SAE-CF joint architecture}

\todo{Describe joint training procedure.}

\section{Controllability mechanism}

The sparse code $\vect{z}_u$ directly exposes interpretable preference knobs.
A user can increase or decrease specific feature activations, and the modified
$\hat{\vect{p}}_u' = \mat{W}_{\text{dec}}\,\vect{z}_u' + \vect{b}_{\text{dec}}$
is passed back to the CF scoring function to re-rank POIs in real time.
