\chapter{Related Work}
\label{chap:related_work}

\todo{Write full related-work chapter.}

\section{Collaborative filtering and matrix factorisation}

Collaborative filtering (CF) exploits the collective behaviour of many users to
generate personalised recommendations.
Matrix factorisation (MF) methods~\cite{koren2009matrix} decompose the
user-item interaction matrix into lower-dimensional latent factor matrices.
Neural CF~\cite{he2017neural} replaces the dot-product interaction with a
multilayer perceptron, further improving accuracy.

\section{Sparse autoencoders}

Autoencoders~\cite{ng2011sparse} learn compressed representations of data
through an encoder--decoder bottleneck.
Imposing an $\ell_1$ or KL-divergence sparsity penalty forces the hidden layer
to represent inputs using only a few active neurons simultaneously.
Recent mechanistic interpretability work~\cite{bricken2023monosemanticity}
has shown that SAEs can decompose dense neural network activations into
monosemantic, human-interpretable dictionary features.

\section{Explainable recommendation}

\todo{Expand this section.}

Despite impressive accuracy, deep recommender models are widely regarded as
black boxes~\cite{zhang2019deep}.
\citeauthor{knijnenburg2012explaining}~\cite{knijnenburg2012explaining} show
that explanations improve user trust and satisfaction.
\citeauthor{jannach2019measuring}~\cite{jannach2019measuring} demonstrate
the business value of explainability.

\section{POI recommendation}

\todo{Expand POI-specific literature.}

POI recommendation~\cite{liu2017experimental} poses unique challenges due to
geographic influence~\cite{ye2011exploiting} and the highly dynamic nature of
user intent.
